Piecewise Regression for Nutritional Outcomes Pre- and Post-CFTR Modulator

### 2022-05-15 StatNet/knot_models.qmd
Update knot_models.qmd

```{r}
#| include: false
# Load libraries, data, etc.
library(DiagrammeR)
library(MultiKink)
library(ggplot2)
data("triceps")
```

@@

Add knots at ages 5, 10, 20, 30, and 40:

```{r}
pred1 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age<5,]))
pred2 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age >=5 & triceps$age<10,]))
pred3 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=10 & triceps$age<20,]))
pred4 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=20 & triceps$age<30,]))
pred5 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=30 & triceps$age<40,]))
pred6 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=40,]))
tri.age.plot + 
  geom_line(data=triceps[triceps$age<5,], 
            aes(y = pred1, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age >=5 & triceps$age<10,], 
            aes(y = pred2, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=10 & triceps$age<20,], 
            aes(y = pred3, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=20 & triceps$age<30,], 
            aes(y = pred4, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=30 & triceps$age<40,], 
            aes(y = pred5, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=40,], 
            aes(y = pred6, x=age), size = 1, col="blue") 
```

@@

```{r}
#| eval: false
grViz("
digraph graph2 {
graph [layout = dot]
# node definitions with substituted label text
node [shape = rectangle]
a [label = '@@1']
b [label = '@@2']
c [label = '@@3']
d [label = '@@4']
e [label = '@@5']
f [label = '@@6']
g [label = '@@7']
h [label = '@@8']
i [label = '@@9']
j [label = '@@10']
k [label = '@@11']
l [label = '@@12']
m [label = '@@13']

>>>>

```{r}
#| include: false
# Load libraries, data, etc.
library(DiagrammeR)
library(MultiKink)
library(ggplot2)
library(broom)
library(knitr)
data("triceps")
```

@@

Add knots at ages 5, 10, and 20:

```{r}
pred1 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age<5,]))
pred2 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age >=5 & triceps$age<10,]))
pred3 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=10 & triceps$age<20,]))
pred4 <- predict(lm(triceps~age, 
                    data = triceps[triceps$age>=20,]))
tri.age.plot + 
  geom_line(data=triceps[triceps$age<5,], 
            aes(y = pred1, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age >=5 & triceps$age<10,], 
            aes(y = pred2, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=10 & triceps$age<20,], 
            aes(y = pred3, x=age), size = 1, col="blue") +
  geom_line(data=triceps[triceps$age>=20,], 
            aes(y = pred4, x=age), size = 1, col="blue") 
```

@@

## Manual coding

```{r}
#| echo: true
# Make new variables
triceps$c5 = (triceps$age-5)*(triceps$age>=5)
triceps$c10 = (triceps$age-10)*(triceps$age>=10)
triceps$c20 = (triceps$age-20)*(triceps$age>=20)
```

```{r}
kable(triceps[c(which(triceps$age<5)[1],which(triceps$age<10&triceps$age>5)[1],
          which(triceps$age>20&triceps$age>10)[1]),],row.names = F)
```

Plot predicted values:

```{r}
# Linear model
mod = lm(triceps ~ age + c5 + c10 + c20,data = triceps)
# Plot
pred = predict(mod)
tri.age.plot +
  geom_line(data=triceps, 
            aes(y = pred, x=age), size = 1, col="blue") 
```

## Model interpretation

```{r}
kable(tidy(mod),digits = 3)
```

The estimates of c5, c10, and c20 indicate change in slope at the respective knot.

```{r}
#| eval: false
grViz("
g [label = '@@7']
h [label = '@@8']
i [label = '@@9']
j [label = '@@10']
k [label = '@@11']
l [label = '@@12']
m [label = '@@13']
